Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-11-11 paper.tex
Updates in discussion

With Boutiques, developers can port their application once and execute
it in several platforms. It removes the technological dependency to a
particular platform and facilitates migration.  Although the
motivating use cases were taken from neuroinformatics, our primary
field of interest, nothing prevents the system from being used in a
variety of other fields.

@@

The \boutiques manifest specification allows describing a wide range
of applications, but it is also getting increasingly complex through
additions such as invocation schemas and dependencies among inputs. In
the future, even more features might be added, for instance to allow
for a better description of when optional outputs are produced, i.e.,
which combination of input values is required. Extending the
\boutiques manifest has two main goals. First, accurately describing
application improves validation: incorrect input values and execution
results are more precisely detected when the application manifest is
comprehensive, which is essential to improve the reliability and
user-friendliness of execution platforms. Secondly, a rich manifest
schema reduces the need for custom wrappers to integrate applications
as even convoluted command lines can be described: this is
particularly interesting for applications installed in containers
since adding a wrapper to the execution stack requires an update in
the container image.

@@

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck to
validation. Using \boutiques, functions can be automatically exported
from frameworks such as Nipype and SPM, creating hundreds of richly
described applications potentially usable by end-users. However, how
to validate that these applications actually execute as expected
remain unclear. A testing framework could be designed and potentially
fed by existing frameworks, to address this issue.

@@

\boutiques is available at \url{https://github.com/boutiques}. We
welcome feedback, issue reporting and pull requests. \boutiques adopts
a bottom-up approach where new features are progressively added based
on feedback from applications and platforms. Beyond the technicalities
discussed in this paper, the availability of a solid core of
applications and platforms in the framework is key to its success,
which we plan to continuously enhance in the future.

>>>>

With Boutiques, developers can port their applications once and execute
them in several platforms. \boutiques removes the technological dependency to a
particular platform and facilitates migration.  Although the
motivating use cases were taken from neuroinformatics, our primary
field of interest, nothing prevents the system from being used in a
variety of other domains.

@@

The \boutiques manifest specification allows describing a wide range
of applications, but it is also getting increasingly complex through
additions such as invocation schemas and dependencies among
inputs. Extending the \boutiques manifest has two main goals. First,
accurately describing applications improves validation: incorrect input
values and execution results are more precisely detected when the
application manifest is comprehensive, which is essential to improve
the reliability and user-friendliness of execution
platforms. Secondly, a rich manifest schema reduces the need for
custom wrappers to integrate applications as even convoluted command
lines can be described: this is particularly interesting for
applications installed in containers since adding a wrapper to the
execution stack requires an update in the container image.

@@

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck
from integration to validation. Using \boutiques, functions can be
automatically exported from frameworks such as Nipype and SPM,
creating hundreds of richly described applications potentially usable
by end-users. However, it remains unclear how to validate that these
applications actually execute as expected. A testing framework could
be designed and potentially fed by existing frameworks, to address
this issue.

@@

\boutiques is available at \url{https://github.com/boutiques}. We
welcome feedback, issue reporting and pull requests. \boutiques adopts
a bottom-up approach where new features are progressively added based
on feedback from applications and platforms. Beyond the technicalities
discussed in this paper, the availability of a solid core of
applications and platforms in the framework is key to its success,
which we plan to continuously enhance.
